DynaProg for Scala

### 2013-01-21 docs/report/benchmarks.tex
Bug int SWat to be fixed, see SWatAffine.

\section{Benchmarks} \label{benchmarks}
In an attempt to provide realistic benchmarks, we tried to gather related implementations. The authors of \cite{gpu_atlp} did not respond to our multiple solicitations. The authors of \cite{swat_mega} were very friendly and provided us their source code. Unfortunately, since they address a different category of problem (they focus on huge serial problems whereas we focus on smaller non-serial problems) their implementation might be biased towards large sequences that our implementation cannot address. Finally, we asked lately the authors of \cite{gpu_rnafold} who did not respond either to our sollicitations.

@@

% ----------------------------------------------
% MatrixMult-512, Mac+JDK7\ul
% Original: 24.937 sec / 24.658 sec
% Optimized concatenation: 19.329 sec / 18.924 sec
% Tabulation as arrays+inline: 14.976 sec / 14.849 sec
\newpage
% ------------------------------------------------------------------------------------------------
{\center\color{red} \noindent\rule{16cm}{0.4pt} \\ XXX: CONTINUE HERE :XXX \\}
% ------------------------------------------------------------------------------------------------
\subsection{Matrix chain multiplication}
Since we have seen that this problem is not input sensitive in (\S\ref{metrics}), we can safely use different random number generators among different implementations without compromising the validity of the results. 
% XXX: is this true ?? For this problem, we modified the hand optimized version because the memory access was slightly different from the generated version (the original hand-optimized version was storing only the matrix multiplication score whereas the generated version also stores the matrix dimension in intermediate results, which proved a better pattern).
Also note that the hand-optimized results are slightly worse than those presented in (\S\ref{baseline_impl}), this is because 64-bit mode is enabled. Because external libraries linked with the Java virtual machine must be in 64 bit, we also enabled this mode in hand-optimized version to maintain a fair comparison, thereby reducing the performance of CUDA operations.
% CPU: DynaProg(Scala), hand optimized C, GAPC and ADPFusion
% GPU: GPU: DynaProg(CUDA), hand optimized CUDA, ATLP and GAPC-OpenCL?

\def\hdr#1#2{\begin{minipage}{3.5cm} {\bf #1} \\[-2pt] \footnotesize #2 \vspace{6pt} \end{minipage}}
\begin{table}[H]\begin{center}{\small\begin{tabular}{llrrrrrrrr}\toprule
& \normalsize\bf Matrix dimension & \normalsize\bf 128 & \normalsize\bf 256 & \normalsize\bf 512 & \normalsize\bf 1024 & \normalsize\bf 2048 & \normalsize\bf 4096 & \normalsize\bf 8192 \\
\midrule \multirow{4}{*}{\rotatebox{90}{\normalsize\bf CPU $\qquad$}}
& \hdr{DynaProg}{Scala version}
	& 			& 			& 			& 			& 			& 			& 		 \\
& \hdr{Optimized}{C, single thread}
	& 0.001		& 0.008		& 0.078		& 1.179		& 19.811		& 206.556	& 2010.485 \\
& \hdr{GAPC}{\cite{gapc_thesis}, C, single thread}
	& 			& 			& 			& 			& 			& 			& 		 \\
& \hdr{ADP Fusion}{\cite{adp_fusion}}
	& 			& 			& 			& 			& 			& 			& 		 \\[-2pt]
\midrule \multirow{4}{*}{\rotatebox{90}{\normalsize\bf GPU $\qquad\quad$}}
& \hdr{DynaProg}{CUDA version}
	& 			& 			& 			& 			& 			& 			& 		 \\
& \hdr{Optimized}{CUDA, 64-bit}
	& 0.011		& 0.020		& 0.074		& 0.315		& 1.651		& 10.354		& 72.376 \\
& \hdr{ATLP}{\cite{gpu_atlp}, rescaled results}
	& 			& 			& 			& 			& 			& 			& 		 \\
& \hdr{GAPC}{\cite{gapc_thesis}, OpenCL}
	& 			& 			& 			& 			& 			& 			& 		 \\
\\[-10pt] \bottomrule\end{tabular}}\end{center}
\caption{Running time of matrix chain multiplication (in seconds)}\end{table}

@@

>>>>

\documentclass[11pt]{article}
\input{inc/style.sty}
\begin{document}

\section{Benchmarks} \label{benchmarks}
In an attempt to provide realistic benchmarks, we tried to gather related implementations. The authors of \cite{gpu_atlp} did not respond to our multiple solicitations. The authors of \cite{swat_mega} were very friendly and provided us their source code. Unfortunately, since they address a different category of problem (they focus on huge serial problems whereas we focus on smaller non-serial problems) their implementation might be biased towards large sequences that our implementation cannot address. Finally, we asked lately the authors of \cite{gpu_rnafold} who did not respond either to our solicitations.

@@

% ----------------------------------------------
% MatrixMult-512, Mac+JDK7\ul
% Original: 24.937 sec / 24.658 sec
% Optimized concatenation: 19.329 sec / 18.924 sec
% Tabulation as arrays+inline: 14.976 sec / 14.849 sec
\newpage
% ------------------------------------------------------------------------------------------------
{\center\color{red} \noindent\rule{16cm}{0.4pt} \\ XXX: CONTINUE HERE :XXX \\}
% ------------------------------------------------------------------------------------------------
\subsection{Matrix chain multiplication}
Since we have seen that this problem is not input sensitive in (\S\ref{metrics}), we can safely use different random number generators among different implementations without compromising the validity of the results.  Also note that the hand-optimized results are slightly worse than those presented in (\S\ref{baseline_impl}), this is because 64-bit mode is enabled. Because external libraries linked with the Java virtual machine must be in 64 bit, we also enabled this mode in hand-optimized version to maintain a fair comparison, thereby reducing the performance of CUDA operations.
% CPU: DynaProg(Scala), hand optimized C, GAPC and ADPFusion
% GPU: GPU: DynaProg(CUDA), hand optimized CUDA, ATLP and GAPC-OpenCL?

\def\hdr#1#2{\begin{minipage}{3.5cm} {\bf #1} \\[-2pt] \footnotesize #2 \vspace{6pt} \end{minipage}}
\begin{table}[H]\begin{center}{\small\begin{tabular}{llrrrrrrrr}\toprule
& \normalsize\bf Matrix dimension & \normalsize\bf 128 & \normalsize\bf 256 & \normalsize\bf 512 & \normalsize\bf 1024 & \normalsize\bf 2048 & \normalsize\bf 4096 & \normalsize\bf 8192 \\
\midrule \multirow{4}{*}{\rotatebox{90}{\normalsize\bf CPU $\qquad$}}
& \hdr{DynaProg}{Scala version}
	& 0.247		& 2.048		& 16.657		& 134.207	& 1101.134	&  			& 		 \\
& \hdr{Optimized}{C, single thread}
	& 0.001		& 0.008		& 0.078		& 1.179		& 19.811		& 206.556	& 2010.485 \\
& \hdr{GAPC}{\cite{gapc_thesis}, C, single thread}
	& 			& 			& 			& 			& 			& 			& 		 \\
& \hdr{ADP Fusion}{\cite{adp_fusion}}
	& 			& 			& 			& 			& 			& 			& 		 \\[-2pt]
\midrule \multirow{4}{*}{\rotatebox{90}{\normalsize\bf GPU $\qquad\quad$}}
& \hdr{DynaProg}{CUDA version}
	& 0.028		& 0.046		& 0.129		& 0.348		& 1.685		& 10.308		& 71.215 \\
& \hdr{Optimized}{CUDA, 64-bit}
	& 0.011		& 0.020		& 0.074		& 0.315		& 1.651		& 10.354		& 72.376 \\
& \hdr{ATLP}{\cite{gpu_atlp}, rescaled results}
	& 			& 			& 			& 			& 			& 			& 		 \\
& \hdr{GAPC}{\cite{gapc_thesis}, OpenCL}
	& 			& 			& 			& 			& 			& 			& 		 \\
\\[-10pt] \bottomrule\end{tabular}}\end{center}
\caption{Running time of matrix chain multiplication (in seconds)}\end{table}

The running time of DynaProg/CUDA includes the overhead of back and forth JNI conversion (scales linearly between 0.018 and 0.057 seconds) but does not include the overhead due to the code generation which decomposes in 0.068 seconds for analysis and code synthesis (once per algebra/grammar pair) and 0.086 + 1.753 seconds for respectively Scala and CUDA compilation (constant time, once per problem dimension).

For DynaProg/Scala we use a variant of the problem description: the original version only stores the matrix multiplication score whereas the modified version also stores the matrix dimension. This allows a speedup of $2.9\times$ probably due to additional function calls to retrieve the values. Also with the default JVM parameters, the program cannot address sequences longer than $\sim 420$ elements due to a stack overflow, for these benchmarks, we increased this limit.

@@

\end{document}
